Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest \\\large{A Bayesian Perspective}

### 2023-02-08 manuscript/manuscript.tex
updated Tables and Graphs

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(CPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support) can improve survival and
diminish anoxic brain injury is a current topic of research. The largest
randomized clinical trial (RCT) examining this question recently
published their results\citep{CPR2023a}. For the primary outcome, 30 day
survival without significant neurological deficit, the authors observed
an odds ratio of 1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR for leading to their conclusion ``In
patients with refractory out-of-hospital cardiac arrest, extracorporeal
CPR and conventional CPR had similar effects on survival with a
favorable neurological outcome''.\citep{CPR2023a}

@@

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial \citep{CPR2023a} and used
for the primary analysis. The ITT analysis has the advantage of
minimizing bias by preserving the prognostic balance afforded by
randomization as well as assuring the validity of the statistical
analyses. ITT assesses subjects based on the group they were initially
(and randomly) allocated to, regardless of whether or not they dropped
out, were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis involves a comparison of treatment groups in a trial that
includes only those patients who completed the treatment they were
originally allocated to. Similarly an ``as treated'' analysis considers
only which treatments subjects received, regardless of their
randomization status and protocol adherence. While both PP and
as-treated analyses alone may lead to bias, in conjunction with an ITT
analysis they may provide additional insights into efficacy and have
also been examined from a Bayesian perspective.

@@

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Using prior
information also underscores the important advantage of Bayesian
analyses to learn sequentially. There were two previous RCTs examining
extracorporeal CPR{[}\citet{RN6759}{]}\citep{RN6751} and while the
protocols are not identical, it may be reasonable to allow this data can
serve as an informed prior which can be updated with the INCEPTION data.
This prior information of the probability of success in each arm,
\(X_i\), can be summarized by a Beta distribution as follow
\[X_i\sim\mbox{Beta}(\alpha_0,\beta_0)\] where in the context of
Bernoulli trials, \(\alpha\) can be interpreted as 1 + number of
successes and \(\beta\) can be interpreted as 1 + number of failures.
The mean and variance of this distribution are
\[\mu=\frac{\alpha}{\alpha+\beta}\]

\[\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}\] where
\(\alpha\) and \(\beta\) are the number of successes and failures
respectively. The difference between the two treatment arms of these
previous studies can then serve as the prior information which in the
Bayesian paradigm can be updated with the INCEPTION data to form a
posterior distribution.

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. We not only calculated the posterior probability
of any benefit (OR \textgreater1.00), but also of clinically meaningful
benefits (defined as OR \textgreater1.10).

All analyses were executed within the integrated development environment
of RStudio and the statistical code can be found on Github
(https://github.com/brophyj/eCPR).

@@

The ITT data from the three pertinent trials is shown in Table 1.
Performing a Bayesian analysis on the INCEPTION trial, using a vague
prior, produces an odds ratio 1.32 (95\% Credible Interval (CrI) 0.54 -
3.22). The closeness of this result to the original analysis confirms
the lack if impact of the vague prior and shows this Bayesian analysis
is dominated by the observed data.

@@

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Table 1 Extracted ITT trial data}\tabularnewline
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endfirsthead
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endhead
INCEPTION & 52 & 56 & 10 & 14 \\
ARREST & 15 & 8 & 0 & 6 \\
PRAGUE & 108 & 86 & 24 & 38 \\
\bottomrule()
\end{longtable}

\hypertarget{bayesian-analysis}{%
\section{Bayesian analysis}\label{bayesian-analysis}}

\begin{longtable}{lrrrc}
\toprule
prior & pt\_est <- ... & llimit <- ... & ulimit <- ... & p.gt.1 <- ... \\ 
\midrule
vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
combined & 1.27 & 0.696 & 2.31 & 0.782 \\ 
enthusiastic & 1.27 & 0.708 & 2.27 & 0.79 \\ 
skeptical & 1.25 & 0.695 & 2.26 & 0.777 \\ 
\bottomrule
\end{longtable}

\hypertarget{summary-of-what-i-dont-understand}{%
\section{\texorpdfstring{\textbf{Summary of what I don't
understand}}{Summary of what I don't understand}}\label{summary-of-what-i-dont-understand}}

This makes only partial sense. 95\% CrI are narrower with informative
priors. OK, but one would think the enthusiastic prior from a trial
ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes
for eCPR vs CPR would shift the HR of 1.32 to the right but instead the
HR falls to have a 1.27. The skeptical prior from PRAGUE which was
stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as
expected

So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)\\
skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)

Tx for enthusiastic prior \textgreater{} Tx skeptical prior (eCPR) but
this seems to be offset by the reverse occurring for the intercept
(CPR).\\
The problem seems to be in how comparable are these previous RCTs as the
baseline success for the standard TX (CRP) is radically different 0/14
vs 24/132.

When I use a prior that uses the enthusiastic intercept prior and the
skeptical Tx prior, I get a bigger downward shift with HR 1.220 95\%
CrI(0.678 2.194) and P\textgreater1 74.5\%

Can you see any other explanation?

Do you have any other insights?

@@

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}

>>>>

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, enhanced, frame hidden]}{\end{tcolorbox}}\fi

@@

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(CPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support (eCPR)) can improve survival
and diminish anoxic brain injury is a current topic of research. The
largest randomized clinical trial (RCT) examining this question recently
published their results\citep{CPR2023a}. For the primary outcome, 30 day
survival without significant neurological deficit, the authors observed
an odds ratio of 1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR for leading to their conclusion ``In
patients with refractory out-of-hospital cardiac arrest, extracorporeal
CPR and conventional CPR had similar effects on survival with a
favorable neurological outcome''.\citep{CPR2023a}

@@

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial \citep{CPR2023a} and used
for the primary analysis. The ITT analysis has the advantage of
minimizing bias by preserving the prognostic balance afforded by
randomization as well as assuring the validity of the statistical
analyses.

@@

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Using prior
information also underscores the important advantage of Bayesian
analyses to learn sequentially. There were two previous RCTs examining
extracorporeal CPR\citep{RN6759, RN6751} and while the protocols are not
identical, it may be reasonable to allow this data to serve as informed
priors for the eCPR parameter, which can be updated with the INCEPTION
data.

This prior information of the probability of eCPR success in each
previous trial, \(X_i\), can be summarized as a normal distribution with
a mean equal to the proportion of successes, \(\hat{p_i}\) with a
standard deviation equal to \[\sqrt{\hat{p_i}*(1-\hat{p_i})}\] As
baseline success rates for standard CPR varies markedly between the
three studies, it was decided to maintain the INCEPTION control baseline
with the vague prior for all analyses and to update only the eCPR arm
with prior information.

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. We not only calculated the posterior probability
of any additional survival with eCPR (OR \textgreater1.00), but also of
clinically meaningful benefits (defined as OR \textgreater1.10).

ITT assesses subjects based on the group they were initially (and
randomly) allocated to, regardless of whether or not they dropped out,
were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis involves a comparison of treatment groups in a trial that
includes only those patients who completed the treatment they were
originally allocated to. Similarly an ``as treated'' analysis considers
only which treatments subjects received, regardless of their
randomization status and protocol adherence. While both PP and
as-treated analyses alone may lead to bias, in conjunction with an ITT
analysis they may provide additional insights into efficacy and have
also been examined from a Bayesian perspective.

All analyses were executed using \texttt{R}\citep{R} within the
integrated development environment of RStudio. Bayesian analyses were
performed using the front end \texttt{rstanarm} package\citep{rstanarm}
to the \texttt{Stan} programming language{[}stan{]}. The statistical
code can be found on Github (https://github.com/brophyj/eCPR).

@@

ITT data from the INCEPTION trial\citep{CPR2023a} and two other
pertinent trials\citep{RN6751, RN6759} that also randomized CPR to eCPR
are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
trial, using the default vague prior, produces an odds ratio (OR) 1.32
(95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms no impact
of the vague prior and shows this Bayesian analysis is completely
dominated by the observed INCEPTION data.

One of the advantages of a Bayesian approach is the ability to make
direct probability statements about the estimand of improved survival
with eCPR. The eCPR probability density function for improved survival
from INCEPTION data with the default vague prior is displayed in Figure
1 and reveals that the probability of enhanced survival with eCPR is
72.72\%. The probability that the improved survival exceeds a 10\%
improvement is 0.66\%.

Another advantage of a Bayesian approach is the possibility of including
previous evidence as informed priors. We considered three possible
scenarios i) using all the available prior RCT
data\citep{RN6751, RN6759} to create a combined prior which was
expressed as a N(0.32, 0.47)\\
ii) using only the ARREST data, labelled an enthusiastic prior, as this
trial was stopped prematurely for efficacy, expressed as a N(0.43,
0.49)\\
iii) using only the PRAGUE data, labelled a skeptical prior, as this
trial was stopped prematurely for futility, expressed as a N(0.31, 0.46)

These different prior probabilities can be updated using the INCEPTION
to create posterior distributions that are summarized in Table 2.
Basically, it is observed that in all cases the point estimate for the
enhanced eCPR survival remains fairly constant but the associated
uncertainty is reduced with the additional data as reflected by the
narrower 95\% CrI. The posterior probability of improved eCPR survival
goes from 80\% with a skeptical prior to 85\% with the enthusiastic
prior.

@@

\setlength{\LTpost}{0mm}
\begin{longtable}{lcccc}
\caption*{
{\large Table 1 Extracted ITT trial data}
} \\ 
\toprule
Trial & Fail CPR (n) & Fail eCPR (n) & Success CPR (n) & Success eCPR (n) \\ 
\midrule
INCEPTION & 52 & 56 & 10 & 14 \\ 
ARREST & 15 & 8 & 0 & 6 \\ 
PRAGUE & 108 & 86 & 24 & 38 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
eCPR = extracorporeal cardiopulmonary resuscitation\\
\end{minipage}

\setlength{\LTpost}{0mm}
\begin{longtable}{lcccc}
\caption*{
{\large Table 2 Probabilities of improved survival with eCPR with various priors}
} \\ 
\toprule
Priors & OR & \multicolumn{2}{c}{95\% CrI} & Probability \\ 
\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-5}
 & point estimate & lower limit & upper limit & p(OR) >1  \\ 
\midrule
Vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
Combined & 1.35 & 0.705 & 2.58 & 0.817 \\ 
Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 \\ 
Skeptical & 1.32 & 0.700 & 2.50 & 0.804 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Vague: default vague prior\\
Combined: prior eCPR data from ARREST + PRAGUE\\
Enthusiastic: prior eCPR data from ARREST alone\\
Skeptical: prior eCPR data from PRAGUE alone\\
\end{minipage}

@@

\includegraphics{manuscript_files/figure-pdf/fig1-1.pdf}

\includegraphics{manuscript_files/figure-pdf/fig2-1.pdf}
